---
title: "R Notebook"
output: html_notebook
---

Loading required files
```{r}
library(caret); 
library(randomForest); 
library(rpart); 
library(doParallel); 
library(e1071); 
library(rpart.plot)
```

Reading train and test files
```{r}
set.seed(12345);
train.data <- read.table(file="/home/ash/pml-training.csv"
                         ,na.strings=c("NA","#DIV/0!",""), header=TRUE, sep=",")
test.data <- read.table(file="/home/ash/pml-testing.csv"
                        ,na.strings=c("NA","#DIV/0!",""),header=TRUE, sep=",")

dim(train.data)
dim(test.data) 

#summary(train.data)
#summary(test.data)

```

Data cleaning : Remove the variables that countains most NAs. 
Remove ‘X’, ‘Name’ and ‘timestamps’ variables as these variables are of no use in prediction
```{r}
train.data <- train.data[, colSums(is.na(train.data)) == 0] 
test.data <- test.data[, colSums(is.na(test.data)) == 0] 

train.data <- train.data[,-c(1,5)]
test.data     <- test.data[,-c(1,5)]


dim(train.data)
dim(test.data) 
```

The train.data set is divided in two parts, one for training and the other for cross validation
The data is split to 3/4 in trainData and testData
```{r}
inTrain <- createDataPartition(train.data$classe, p=3/4, list=FALSE)
trainData <- train.data[inTrain, ]
testData <- train.data[-inTrain, ]

dim(trainData)
dim(testData) 
```

Using Random forest algorithm for creating model
trainControl is the function to for training model
```{r}

train_model <- trainControl(method="cv", 5)
model_RF <- train(classe ~ ., data=trainData, method="rf", trControl=train_model, ntree=250)
model_RF

```

Performance evaluation of the Random forest algorithm.
Now we want to estimate the performance of the model on the validation data set.
A confusionMatrix function and both the estimated accuracy and the the estimated
out-of-sample error of the model are calculated.

```{r}
predictRf <- predict(model_RF, testData)
confusionMatrix(testData$classe, predictRf)

```

```{r}
accuracy <- postResample(predictRf, testData$classe)
accuracy
```
RF Model accuracy is 0.9995922, almost 99.9%


The out of sample error for RF model is shown below
```{r}
outsamerr <- (1 - as.numeric(confusionMatrix(testData$classe, predictRf)$overall[1]))
outsamerr
```

Prediction for the Test Data Set (testRaw).
Prediction using the model modelRF just built using the raw initial data testRaw

```{r}
final_result <- predict(model_RF, test.data)

final_result

```

```{r}
tree_tree <- rpart(classe ~ ., data=trainData, method="class")
prp(tree_tree) 
```

